A Formalization of Complete Discrete Valuation Rings and Local Fields

### 2023-09-09 CPP_paper/CPP_paper.tex
get started on section 2.1

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
%\documentclass[sigplan,screen]{acmart}
\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

@@

\section{Introduction}
\begin{itemize}
\item Say all rings are commutative and unitary.
\item Remember to create the graph
\item Acknowledge de Frutos-Fernández and Yaël Dillies for allowing us to use their work.
\end{itemize}
\section {Discrete Valuation Rings}\label{section:dvr}
\mi{
\subsection{Valuations}\label{subsection:valuations}
\begin{itemize}
	\item Properties of valuations valued in $\ZZ_{m0}$, together with the \code{multiplicative} and \texttt{with\_zero} classes, and clashes with "usual" intuition. Keep in mind that \texttt{with\_zero multiplicative} $\ZZ$ is "the same" as $q^{-\ZZ}\cup \{0\}$.
	\item Difference between \texttt{valued} (which is a class, and carries a topology/uniformity) and \texttt{valuation} that is just a function.
	\item We put \texttt{valued} instances on fields but "never" on rings that are not fields.
	\item There is something called \texttt{valuation\_ring} that has (almost) nothing to do; but \texttt{valuation\_subring} has.
	\item Comment about the choice of \texttt{.integers} instead of \texttt{.valuation\_subring}
\end{itemize}
}
\subsection{Definition of Discrete Valuation Rings}\label{subsection:def_dvr}
Let $R$ be a ring, \fil{as above assumed} commutative and unitary. We begin by recalling a classical result:
\begin{theorem}[see\dots]\label{thm:dvr_tfae}
The following properties are equivalent:
\begin{listResults}
\item foo
\item foo
\item foo
\end{listResults}
\end{theorem}
A ring satisfying the equivalent properties of Theorem~\ref{thm:dvr_tfae} is called a discrete valuation ring (often shortened as DVR). The definition is motivated by the fact that a DVR $R$ can always be endowed with a class of valuations, 
\[
v\colon R\longrightarrow \ZZ\cup\{+\infty\}\fil{\text{ or use the mathlib convention?}}.
\]
For each of these valuations, the image $\image(v)\subseteq\ZZ\cup \{+\infty\}$ is the free monoid generated by $v(r)$, where $r$ is any generator of the maximal ideal $\maxid$ of $R$. Upon replacing $v$ by $v/v(r)$, for some generator $r$ of $\maxid$, we can assume that this free monoid coincides with $\NN$: in this case the valuation is said to be normalised, and the elements of valuation equal to $1$ are called uniformizers. One basic result is that, for a normalised valuation, the uniformisers are exactly the set of generators of $\maxid$, \fil{because \dots}. Moreover, there exists a unique normalised valuation on a DVR, since every valuation is uniquely determined by its value on the generators of $\maxid$. We denote it by ${\normalised}_R$, or simply by $\normalised$ when the DVR $R$ can be understood.

>>>>

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
%\documentclass[sigplan,screen]{acmart}
%\documentclass[sigplan,10pt,anonymous,screen]{acmart}
\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

@@

\section{Introduction}
\begin{itemize}
\item Say all rings are commutative and unitary.
\item Remember to create the graph
\item Acknowledge de Frutos-Fernández and Yaël Dillies for allowing us to use their work.
\item Add  \href{https://github.com/mariainesdff/local_class_field_theory}{link to repository}. \mi{Links are rendered in black with the "review" option, which means they will probably be missed. In the sphere eversion paper they have a symbol next to each link; do you know how to achieve this? Otherwise we could ask them.}
\end{itemize}
\section {Discrete Valuation Rings}\label{section:dvr}

\subsection{Valuations}\label{subsection:valuations}
Let $R$ be a ring. A \textit{valuation} $v$ on $R$ is a map $v : R \to \Gamma_0$  to a linearly ordered commutative monoid with zero $\Gamma_0$ such that
\begin{enumerate}
	\item $v (0) = 0$,
	\item $v (1) = 1$,
	\item $v (x \cdot y) = v (x) \cdot v (y)$ for all $x, y \in R$, and
	\item $v (x + y) \le \max {v (x), v (y)}$ for all $x, y \in R$.
\end{enumerate}

For instance, when $R$ is a Dedekind domain of Krull dimension 1 with fraction field $K$ and $\m$ is a maximal ideal of $R$, we can define a function $c_\m : R \to \ZZ \cup \{ \infty\}$ by sending zero to $\infty$ and any nonzero $r \in R$ to the number of times that $\m$ appears in the factorization of the principal ideal $(r)$. Note that this number is well defined because ideals in a Dedekind domain admit a unique factorization into maximal ideals, up to reordering. We can extend the function $c_\m$ to $K$ by setting $c_\m (r/s) := c_\m (r) - c_\m (s)$ for any $r, s \in R$ with $s \ne 0$.

Then, for any real number $n_\m > 1$, it is easy to check that the function
$ v_\m: K \to n_\m^{\mathbb{Z} \cup \{-\infty\}} = n_\m^\mathbb{Z} \cup \{0\}$ sending $r$ to $n_\m^{-\text{val}_v(r)}$ is a valuation on $K$, which we call the \textit{$\m$-adic valuation}. For example, when $R = \ZZ$, $K = \QQ$ and $\m = (p)$ is the ideal generated by a prime number $p$, $v_p$ is the $p$-adic (multiplicative) valuation on $\QQ$, also known as the $p$-adic absolute value.

We will represent valuations in Lean using the \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/basic.html#valuation}{valuation} structure available in mathlib, which agrees with our above definition. A valuation on a ring $R$ induces a topology on $R$ for which the ring operations are continuous, hence turning $R$ into a topological ring.

Mathlib also provides the class \href{https://leanprover-community.github.io/mathlib_docs/topology/algebra/valuation.html#valued}{valued}, which bundles together a ring with a distinguished valuation and the uniform space structure induced by this valuation. This class is designed for rings in which there is a canonical valuation. For example, on the rational number $\QQ$ there are

\mi{
\begin{itemize}
	\item Properties of valuations valued in $\ZZ_{m0}$, together with the \code{multiplicative} and \texttt{with\_zero} classes, and clashes with "usual" intuition. Keep in mind that \texttt{with\_zero multiplicative} $\ZZ$ is "the same" as $q^{-\ZZ}\cup \{0\}$.
	\item Difference between \texttt{valued} (which is a class, and carries a topology/uniformity) and \texttt{valuation} that is just a function.
	\item We put \texttt{valued} instances on fields but "never" on rings that are not fields.
	\item There is something called \texttt{valuation\_ring} that has (almost) nothing to do; but \texttt{valuation\_subring} has.
	\item Comment about the choice of \texttt{.integers} instead of \texttt{.valuation\_subring}
\end{itemize}
}
\subsection{Definition of Discrete Valuation Rings}\label{subsection:def_dvr}
Let $R$ be a ring, \fil{as above assumed} commutative and unitary. We begin by recalling a classical result:
\begin{theorem}[see\dots]\label{thm:dvr_tfae}
The following properties are equivalent:
\begin{listResults}
\item foo
\item foo
\item foo
\end{listResults}
\end{theorem}
A ring satisfying the equivalent properties of Theorem~\ref{thm:dvr_tfae} is called a discrete valuation ring (often shortened as DVR). The definition is motivated by the fact that a DVR $R$ can always be endowed with a class of valuations, 
\[
v\colon R\longrightarrow \ZZ\cup\{+\infty\}\fil{\text{ or use the mathlib convention?}}.
\]
For each of these valuations, the image $\image(v)\subseteq\ZZ\cup \{+\infty\}$ is the free monoid generated by $v(r)$, where $r$ is any generator of the maximal ideal $\maxid$ of $R$. Upon replacing $v$ by $v/v(r)$, for some generator $r$ of $\maxid$, we can assume that this free monoid coincides with $\NN$: in this case the valuation is said to be normalised, and the elements of valuation equal to $1$ are called uniformizers. One basic result is that, for a normalised valuation, the uniformisers are exactly the set of generators of $\maxid$, \fil{because \dots}. Moreover, there exists a unique normalised valuation on a DVR, since every valuation is uniquely determined by its value on the generators of $\maxid$. We denote it by ${\normalised}_R$, or simply by $\normalised$ when the DVR $R$ can be understood.
